Learning STRIPS action models with classical planning

### 2017-10-12 doc/icaps-18/strips-learning.tex
added more content to the ICAPS18 paper

\maketitle
\begin{abstract}
This paper presents a novel approach for learning STRIPS action models from examples that compiles this inductive learning task into a classical planning task. Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed. What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can also be used to validate whether an observed plan follows a given STRIPS action model. 
\end{abstract}

\section{Introduction}
Besides plan synthesis, planning action models are also useful for plan/goal recognition~\cite{ramirez2010probabilistic}. In both tasks, off-the-shelf planners require reasoning about action models that correctly and completely capture the possible world transitions~\cite{ghallab2004automated,geffner:book:2013}. Unfortunately, building such planning action models is complex, even for planning experts, and this knowledge acquisition bottleneck limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.

On the other hand, Machine Learning (ML) techniques are able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}. The application of inductive ML techniques for learning planning action models is not straightforward though:
\begin{itemize}
\item The inputs to ML algorithms usually are finite vectors encoding the value of fixed features for a given set of objects. The input for learning planning action models traditionally are sets of observations of plan executions (each with a possibly different length).
\item The traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks). When learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions of a given planning domain. 
\end{itemize}

Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the intermediate states traversed by the example plans. Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation so an off-the-shelf planner can be used to address the inductive learning task. In addition the compilation approach presents the following benefits:
\begin{enumerate}
\item It is flexible to different amounts of available input knowledge. The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
\item Can exploit previous knowledge about the action model in the form of partially specified action models.
\item Can be used to validate whether an observed plan follows a given STRIPS action model.
\end{enumerate}

The paper is organized as follows. The first section presents the classical planning model and its extension to conditional effects (a requirement of the proposed compilation). The second section formalizes the task of learning STRIPS action models from observations of plan executions with regard to different amounts of input knowledge. The third section describes our approach for addressing these inductive learning tasks by compiling them into classical planning. The fourth section evaluates the performance of our approach. Finally, the last section discusses the strengths and weaknesses of our approach and proposes some future work.

@@

\subsection{Classical planning}
We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent) and let $\neg L=\{\neg l:l\in L\}$ be the complement of $L$. We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.

A {\em state} $s$ is then a total assignment of values to fluents, i.e. $|s|=|F|$, so the size of the state space $2^{|F|}$. Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions, but we often abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of positive effects $\eff^+(a)\in\mathcal{L}(F)$, and a set of negative effects $\eff^-(a)\in\mathcal{L}(F)$. An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff^-(a))\cup\eff^+(a)$.

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The plan $\pi$ {\em solves} $P$ if and only if $G\subseteq s_n$, i.e.~if the goal condition is satisfied following the application of $\pi$ in $I$.

In this work we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}. There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$. Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.

Likewise, we assume that each action in $A$ is instantiated from an STRIPS operator schema. Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.
\begin{figure}[hbt]
\begin{footnotesize}
\begin{verbatim}
(:action stack
  :parameters (?x1 ?x2)
  :precondition (and (holding ?x1) 
                     (clear ?x2))
  :effect (and (not (holding ?x1)) 
               (not (clear ?x2))
               (clear ?x1) 
               (handempty) 
               (on ?x1 ?x2)))
\end{verbatim}
\end{footnotesize}
 \caption{\small Example of a STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.}
\label{fig:stack}
\end{figure}

Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names}. $|\Omega_v|$ is given by the maximum arity of an action in the given planning frame. For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).

Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} $\Omega_v$. In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.

We are now ready to define a STRIPS operator schema as a tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$:
\begin{itemize}
\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents an operator {\em header} defined by its corresponding action name and a enumeration of the variable names, $pars(\xi)\in\Omega_v^{ar(\xi)}$. The headers for a 4-operator blocksworld are then {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
\item The preconditions $pre(\xi)\subseteq F_v$, the positive effects $add(\xi)\subseteq F_v$, and the negative effects $del(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}

\subsection{Classical planning with conditional effects}
Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects. We use conditional effects because they allow us to compactly define actions whose particular effects depend on the current state. Many classical planners cope with conditional effects without compiling them away. In fact, the support of PDDL conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015}.

@@

@@

\subsection{Learning from labeled plans}
This learning task is formalized as $\Lambda=\tup{\Psi,\Pi,\Sigma}$: 
\begin{itemize}
\item $\Psi$, the set of predicates that define the abstract state space of a given planning domain. 
\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans s.t. each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces a state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, a set of labels s.t. each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has an associated label $\sigma_t=(s_0^t,s_{n}^t)$ where $s_{n}^t$ is the state resulting from executing the plan $\pi_t$ starting from the state $s_0^t$. 
\end{itemize}

To illustrate this, Figure~\ref{fig:lexample}, shows the contain of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$ for a learning task $\Lambda$ with a single learning example from the blocksworld for inverting a tower of four blocks.

@@

\begin{subfigure}{.3\textwidth}
\begin{small}
{\tt ;;; Plan $\pi_1$}
\begin{verbatim}
0: (unstack A B)
1: (putdown A)
2: (unstack B C)
3: (stack B A)
4: (unstack C D)
5: (stack C B)
6: (pickup D)
7: (stack D C)
\end{verbatim}
\end{small}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
{\small\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
\begin{lstlisting}[mathescape]
\end{lstlisting}
\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
\node (initD) [draw]{D};
\node (initC) [draw, above=of initD.north]{C};
\node (initB) [draw, above=of initC.north]{B};
\node (initA) [draw, above=of initB.north]{A};
\draw (-1,-0.25) -- (2.5,-0.25);

@@

 \caption{\small Example of a learning task with a single learning example that corresponds to invert a tower of four blocks.}
\label{fig:lexample}
\end{figure}

A solution to the learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$. With this regard, the operator schema shown in Figure~\label{fig:stack} is compliant with the learning task illustrated by the Figure~\ref{fig:lexample}.

\subsection{Learning from initial/final states}
Here we reduce the amount of input knowledge provided to the learning task. Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e. $\Pi'$ that does not contain a set of plans but the number of actions of each plan and the headers of the operators schema. The learning task is hence redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$. While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the result of its plan execution but knowing the number of actions performed by the agent. In the case of the learning example shown in Figure~\ref{fig:lexample}, $|\pi_1|=8$.

@@

In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only synthesize a possible action models but also the actions could have produced the given labels (this information about the actions is no longer given in the learning examples).

\section{Learning STRIPS action models with planning}
Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects $P_{\Lambda}$, $P_{\Lambda'}$ or $P_{\Lambda''}$. The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item Programs the STRIPS action model. For each $\xi\in\Xi$, determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
\item Validates the programmed operator schema $\Xi$ with the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$.  $\Xi$ is used, at every {\small $1\leq t\leq \tau$}, to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$. If information about the plans is given, then it is used to constrain the validation of the programmed opertator schema $\Xi$.
\end{enumerate}

Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ class. In particular the classical plan programs and validates the operator schema {\tt\small stack} using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}. The details of the compilation are given in the remaining of the section.

\begin{figure}[hbt]
\begin{footnotesize}
\begin{verbatim}
     0 : (program_pre_clear_stack_var1)
     1 : (program_pre_handempty_stack_)
     2 : (program_pre_holding_stack_var2)
     3 : (program_pre_on_stack_var1_var1)
     4 : (program_pre_on_stack_var1_var2)
     5 : (program_pre_on_stack_var2_var1)
     6 : (program_pre_on_stack_var2_var2)
     7 : (program_pre_ontable_stack_var1)
     8 : (program_pre_ontable_stack_var2)
     9 : (program_eff_clear_stack_var1)
    10 : (program_eff_clear_stack_var2)
    11 : (program_eff_handempty_stack_)
    12 : (program_eff_holding_stack_var1)
    13 : (program_eff_on_stack_var1_var2)
    14 : (unstack a b i1 i2)
    15 : (putdown a i2 i3)
    16 : (unstack b c i3 i4)
    17 : (stack b a i4 i5)
    18 : (unstack c d i5 i6)
    19 : (stack c b i6 i7)
    20 : (pickup d i7 i8)
    21 : (stack d c i8 i9)
    22 : (validate_1)
\end{verbatim}
\end{footnotesize}
 \caption{\small Example of a plan for programing and validating the operator schema unstack using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.}
\label{fig:plan}
\end{figure}

To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, one for each leaning example, that belong to the same planning frame (i.e. same fluents and actions and differ only in the initial state and goals). The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of objects appearing in the examples. Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state. The set of actions, $A=\emptyset$, is empty because the action model is unknown. Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.

Now we are ready to formalize our compilations for learning STRIPS action models with classical planning. We start with $\Lambda''$ because it is the learning task that requires the smallest amount of input knowledge and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.

>>>>

\maketitle
\begin{abstract}
This paper presents a novel approach for learning STRIPS action models from examples that compiles this inductive learning task into a classical planning task. Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are observed. What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can also be used to validate whether observation of plan executions follow a given STRIPS action model. 
\end{abstract}

\section{Introduction}
Besides plan synthesis~\cite{ghallab2004automated,geffner:book:2013}, planning action models are also useful for plan/goal recognition~\cite{ramirez2012plan}. In both tasks, off-the-shelf planners require reasoning about action models that correctly and completely capture the possible world transitions. Unfortunately, building such planning action models is complex, even for planning experts, and this knowledge acquisition bottleneck limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.

On the other hand, Machine Learning (ML) techniques have shown able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}. The application of inductive ML techniques for learning planning STRIPS action models from examples is not straightforward though:
\begin{itemize}
\item The inputs to ML algorithms usually are finite vectors that encode the value of fixed features for a given set of objects. The input for learning planning action models traditionally are sets of observations of plan executions (each with possibly a different length).
\item The traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks). When learning STRIPS action models, the output is not a scalar but the sets of preconditions and effects of each action that define the possible state transitions of a given planning domain. 
\end{itemize}

Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans. Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation (so an off-the-shelf planner can be used to address the inductive learning task). In addition, the compilation approach presents the following benefits:
\begin{enumerate}
\item It is flexible to different amounts of available input knowledge. The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
\item Can exploit previous knowledge about the action model in the form of partially specified action models.
\item Can be used to validate whether observed plan executions are valid for a given STRIPS action model.
\end{enumerate}

The paper is organized as follows. The first section presents the classical planning model, its extension to conditional effects (a requirement for the proposed compilation) and the STRIPS action model. The second section formalizes the task of learning STRIPS action models from observations of plan executions with regard to different amounts of input knowledge. The third section describes our approach for addressing these inductive learning tasks by compiling them into classical planning. The fourth section reports the collected data during the evaluation of the performance of our approach. Finally, the last section discusses the strengths and weaknesses of our approach and proposes some future work.

@@

\subsection{Classical planning}
We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent) and let $\neg L=\{\neg l:l\in L\}$ be the complement of $L$. We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.

A {\em state} $s$ is then a total assignment of values to fluents, i.e. $|s|=|F|$, so the size of the state space is $2^{|F|}$. Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of negative effects $\eff^-(a)\in\mathcal{L}(F)$, and a set of positive effects $\eff^+(a)\in\mathcal{L}(F)$. An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff^-(a))\cup\eff^+(a)$.

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The plan $\pi$ {\em solves} $P$ if and only if $G\subseteq s_n$, i.e.~if the goal condition is satisfied in the state reached after following the application of $\pi$ in $I$.

\subsection{Classical planning with conditional effects}
Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects. We use conditional effects because they allow us to compactly define actions whose particular effects depend on the current state. Many classical planners cope with conditional effects without compiling them away. In fact, the support of conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015}.

@@

\subsection{The STRIPS action model}
In this work we address the task of learning  action  schema  in  the  STRIPS language~\cite{fikes1971strips}. With this regard, we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}. There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$. Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.

Likewise, we assume that each action in $A$ is instantiated from a STRIPS operator schema. Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld~\cite{slaney2001blocks} represented in PDDL.
\begin{figure}[hbt]
\begin{footnotesize}
\begin{verbatim}
(:action stack
  :parameters (?x1 ?x2 - object)
  :precondition (and (holding ?x1) 
                     (clear ?x2))
  :effect (and (not (holding ?x1)) 
               (not (clear ?x2))
               (clear ?x1) 
               (handempty) 
               (on ?x1 ?x2)))
\end{verbatim}
\end{footnotesize}
 \caption{\small Example of a STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.}
\label{fig:stack}
\end{figure}

Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names}. The cardinality of this set, $|\Omega_v|$, is given by the maximum arity of an action in the given planning frame. For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).

Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} in $\Omega_v$. In the blocksworld this set is defined as $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.

We are now ready to define a STRIPS action model $\Xi$ as a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$:
\begin{itemize}
\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents the operator {\em header} defined by its corresponding action name and a enumeration of the variable names given by the arity of the operator, $pars(\xi)\in\Omega_v^{ar(\xi)}$. The headers for a 4-operator blocksworld are then {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}

@@

\subsection{Learning from labeled plans}
This learning task is formalized as $\Lambda=\tup{\Psi,\Pi,\Sigma}$: 
\begin{itemize}
\item $\Psi$, the set of predicates that define the abstract state space of a given planning domain. 
\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans s.t. each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces a state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, is a set of labels s.t. each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has an associated label $\sigma_t=(s_0^t,s_{n}^t)$ where $s_{n}^t$ is the state resulting from executing the plan $\pi_t$ starting from the state $s_0^t$. 
\end{itemize}

To illustrate this, Figure~\ref{fig:lexample}, shows the contain of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$ for a $\Lambda$ task that aims learning the STRIPS action model of the blocksworld. The shown learning task has a single learning example that corresponds to observing a plan execution for inverting a tower of four blocks.

@@

\begin{subfigure}{.3\textwidth}
\begin{small}
{\tt ;;; Plan $\pi_1$}
\begin{verbatim}
0: (unstack A B)
1: (putdown A)
2: (unstack B C)
3: (stack B A)
4: (unstack C D)
5: (stack C B)
6: (pickup D)
7: (stack D C)
\end{verbatim}
\end{small}
\end{subfigure}%
\begin{subfigure}{.3\textwidth}
{\small\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
\begin{lstlisting}[mathescape]
\end{lstlisting}
\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
\node (initD) [draw]{D};
\node (initC) [draw, above=of initD.north]{C};
\node (initB) [draw, above=of initC.north]{B};
\node (initA) [draw, above=of initB.north]{A};
\draw[thick] (-1,-0.25) -- (2.5,-0.25);

@@

 \caption{\small Learning task in the blocksworld with a single learning example that corresponds to inverting a tower of four blocks.}
\label{fig:lexample}
\end{figure}

A solution to the learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$. With this regard, the operator schema shown in Figure~\ref{fig:stack} is compliant with the learning task illustrated by the Figure~\ref{fig:lexample}.

\subsection{Learning from initial/final states}
Here we reduce the amount of input knowledge provided to the learning task. Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e. $\Pi'$ that does not contain a set of plans but the number of actions of each plan. The learning task is hence redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$. While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the result of its plan execution but knowing the number of actions performed by the agent. In the case of the learning example shown in Figure~\ref{fig:lexample}, $|\pi_1|=8$.

@@

In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only synthesize a possible action models but also the actions could have produced the given labels (this information about the actions is no longer given in the learning examples). However, we assume that at the headers of the operators schema are known.

\section{Learning STRIPS action models with planning}
Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects $P_{\Lambda}$, $P_{\Lambda'}$ or $P_{\Lambda''}$. The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
\begin{enumerate}
\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
\item Validates the programmed the STRIPS action model $\Xi$ using the available input knowledge ($\Pi$, if available, and the given set of labels $\Sigma$).  At every learning example {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$. If information about the corresponding plan $\pi_t$ is available, then it is used to constrain the validation of the programmed STRIPS action model $\Xi$.
\end{enumerate}

Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ class. The shown classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}. In particular the steps $[0,8]$ are actions for programming the preconditions of the operator, steps $[9,13]$ are actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operator. The details of the compilation are given in the remaining of the section.

\begin{figure}[hbt]
\begin{footnotesize}
\begin{verbatim}
     0 : (program_pre_clear_stack_var1)
     1 : (program_pre_handempty_stack_)
     2 : (program_pre_holding_stack_var2)
     3 : (program_pre_on_stack_var1_var1)
     4 : (program_pre_on_stack_var1_var2)
     5 : (program_pre_on_stack_var2_var1)
     6 : (program_pre_on_stack_var2_var2)
     7 : (program_pre_ontable_stack_var1)
     8 : (program_pre_ontable_stack_var2)
     9 : (program_eff_clear_stack_var1)
    10 : (program_eff_clear_stack_var2)
    11 : (program_eff_handempty_stack_)
    12 : (program_eff_holding_stack_var1)
    13 : (program_eff_on_stack_var1_var2)
    14 : (apply_unstack a b i1 i2)
    15 : (apply_putdown a i2 i3)
    16 : (apply_unstack b c i3 i4)
    17 : (apply_stack b a i4 i5)
    18 : (apply_unstack c d i5 i6)
    19 : (apply_stack c b i6 i7)
    20 : (apply_pickup d i7 i8)
    21 : (apply_stack d c i8 i9)
    22 : (validate_1)
\end{verbatim}
\end{footnotesize}
 \caption{\small Example of a plan for programing and validating the operator schema unstack using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.}
\label{fig:plan}
\end{figure}

To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, one for each leaning example, that belong to the same planning frame (i.e. same fluents and actions and differ only in the initial state and goals). The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of different objects appearing in the examples. Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state. The set of actions, $A=\emptyset$, is empty because the action model is unknown. Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.

Now we are ready to formalize our compilations for learning STRIPS action models with classical planning. We start with $\Lambda''$ because it is the learning task that requires the smallest amount of input knowledge, and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.

Given a learning task $\Lambda''=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda''}=\tup{F_{\Lambda''},A_{\Lambda''},I_{\Lambda''},G_{\Lambda''}}$ such that:
\begin{itemize}
\item $F_{\Lambda''}$ extends $F$ with:
\begin{itemize}
\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$). If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$. 
\item A fluent $mode_{prog}$ indicating whether the operator schema are being programmed or they are already programmed operators so are being validated in the learning examples. 
\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
\end{itemize}
\item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e. the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially, any operator schema, contains all the possible preconditions, no negative effect and no positive effect) and $mode_{prog}$. 
\item $G_{\Lambda''}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
\item $A_{\Lambda''}$ contains actions of three different kinds:
\begin{enumerate}
\item The actions for programming the operator schema. This set includes:
\begin{itemize}
\item The actions for removing a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
